DATA 621 Homework #3

### 2019-10-30 hw3/HW3.Rmd
Update HW3.Rmd
Added some description

### Are There Missing Values?

@@

### Splitting the Data

```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Exploratory Data Analysis

@@

The following show the how predictors are distributed between the areas with crime rates higher than the median (blue) and below the median (red).  What we are looking for is variables that could split data into two groups.

@@

NOX seems to be the best variable to divide the data into the two groups

@@

Applying occam's razor we will create a baseline that only has one predictor.  Any model will have to out preform this simplest model.

@@

Now we will try adding all other variables and work backwards to eliminate non-significant predictors:

```{r}
fullmodel <- glm(target ~ nox + dis, family = binomial(link = "logit"), train)
summary(fullmodel)
test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
results <- rbind(results,tibble(model = "fullmodel",
                                predictors = 12,F1 = cm$byClass[7],
                                deviance=fullmodel$deviance, 
                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
                                aic=fullmodel$aic))
cm
```

The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, the new variable doesn't seem to be significant.

@@

Looking at our models together, `model1` looks like the best so far based on highest F1 score and lowest deviance & AIC.

@@

## Model Selection

>>>>

### Are There Missing Values?
First we look at the data to see if any variables have missing data:

@@

### Splitting the Data
Next, we look to split our data between a training set (`train`) and a test data set (`test`). We'll use a 70-30 split between train and test, respectively.
```{r}
set.seed(42)
train_index <- createDataPartition(df$target, p = .7, list = FALSE, times = 1)
train <- df[train_index,]
test <- df[-train_index,]
```

### Exploratory Data Analysis
Now, let's look at our training data. By looking at a correlation matrix, we can see which variables may be too correlated to be included together in a model as predictor variables. This will help us later during the model selection process.

@@

Next we will look at each potential predictor and how it is distributed across the target variable.

The following plots show how predictors are distributed between a positive target variable (areas with crime rates higher than the median, i.e. blue) and a negative target variable (areas with crime rates below the median, i.e. red). What we are looking for is variables that show way to split data into two groups.

@@

Looking at the plots above, the `nox` variable seems to be the best variable to divide the data into the two groups.

@@

We start by applying occam's razor and create a baseline model that only has one predictor. Any model we build beyond that will have to outperform this simplest model.

@@

Next we try adding every other variable, to build a full model. From here we can work backwards and eliminate non-significant predictors:

```{r}
fullmodel <- glm(target ~ ., family = binomial(link = "logit"), train)
summary(fullmodel)
test$fullmodel <- ifelse(predict.glm(fullmodel, test,"response") < 0.5, 0, 1)
cm <- confusionMatrix(factor(test$fullmodel), factor(test$target),"1")
results <- rbind(results,tibble(model = "fullmodel",
                                predictors = 12,F1 = cm$byClass[7],
                                deviance=fullmodel$deviance, 
                                r2 = 1-fullmodel$deviance/fullmodel$null.deviance,
                                aic=fullmodel$aic))
cm
```

The full model has an F1 score is `r cm$byClass[7]`, which is a bit higher than before. However, many variables do not seem to be significant.

@@

Since the assignment mentions that the purpose is prediction, we will prefer F1 score as our measure of model success.

@@

Looking these three models together, `model1` looks like the best one. Although the full model scored an F1 that was ever-so-slightly higher on our test data set, `model1` has half the predictors and it's scores are almost exactly the same as `fullmodel`.
